---
title: "Predictive Modeling of Correct Barbell Lifts"
author: "skywa"
date: "Friday, February 20, 2015"
output: html_document
---

Various wearable devices collect information about personal physical activity. These often quantify how much activity is done, but do not predict how well they do it. This study uses a prediction algorithm to use the accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict if they performed barbell lifts correctly or incorrectly in 5 different ways.

First, let's install two libraries: randomForest and caret
```{r}
# install.packages("randomForest")
# install.packages("caret")
library(randomForest); library(caret)
```

Import the data
```{r, cache=TRUE}
training <- read.csv("pml-training.csv", header=T, na.strings=c(""," ",NA))
testing <- read.csv("pml-testing.csv", header=T, na.strings=c(""," ",NA))
# str(training)
```

There are quite a few columns that are almost all NAs. These are removed for both the training and the test set. All but the 'classe' variable are converted to a numeric class.
```{r}
train <- training[, !apply(training, 2, function(x) any(is.na(x)))]
test <- testing[, !apply(training, 2, function(x) any(is.na(x)))]
train <- train[,-c(1:7)]
test <- test[,-c(1:7)]
newtrain <- cbind(data.frame(sapply(train[,-53], as.numeric), train[,53]))
newtest <- cbind(data.frame(sapply(test[,-53], as.numeric), test[,53]))
colnames(newtrain)[53] <- "classe";  colnames(newtest)[53] <- "predict_id"
```

To complete some cross-validation, the training set is split into training and testing components (intrain and intest).
```{r}
set.seed(941)
indexTrain <- createDataPartition(y=train$classe, p=0.6, list=FALSE)
intrain <- newtrain[indexTrain,]
intest <- newtrain[-indexTrain,]
```

Using randomForest, a model is created to predict 'classe' using the intrain data.
```{r, cache=T}
set.seed(25703)
rf <- randomForest(formula = classe ~ ., data=intrain, ntree=500)
rf
```

Testing the model on the 'intest' data, we get high (>99%) accuracy, shown on the confusion matrix.
```{r}
pred <- predict(rf, intest)
cf <- confusionMatrix(pred, intest$classe)
cf
```

The random forest model is used on the original test data set to predict the A, B, C, D, and E for the barbell lifts.
```{r}
FinalPred <- predict(rf, newdata=newtest)
FinalPred
```

Files were created to submit as answers.
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i], file=filename, quote=FALSE, row.names=FALSE, col.names=FALSE)}}

pml_write_files(FinalPred)
```

All the predictions were correct!